Product graphics

### 2009-12-05 product-graphics.tex
Initial work on variations

These techniques can also be used with continuous data, if the data is binned to create a categorical variable. This gives rise to the histogram and spinogram, analogues of the bar chart and spine chart respectively. A long standing tradition is that no gaps are displayed between adjacent rectangles when used for originally continuous data. Two examples are shown in Figure \ref{fig:cont-examples}.

@@

\subsection{Nested data}
\label{sub:nested_data}

We distinguish nested data from crossed data by the number of interactions with missing values.  Completely crossed data contains every combination of all variables.  Nested data contains

Aka sparse vs. dense.

It is not always possible to determine nested data from inspection of the data alone, but may need knowledge of the data collection.  For example, if a survey of teachers within schools was performed, and each teacher was identified by a identifier unique within a school, it might look like the data was only missing a few combinations, when in fact it is meaningless.

Or when you have a hierarchy.

Example dataset - Doug Bates teachers/schools?

Possibilities: create new variable which is interaction of nested vars, or drop missings/zeros.  Treemaps deal almost exclusively with nested data.

\section{Data analysis} % (fold)
\label{sec:data_analysis}

% section data_analysis (end)

>>>>

These techniques can also be used with continuous data, if the data is binned to create a categorical variable. There are many different possible choices of binning algorithms, but we will focus on two:

\begin{itemize}
  \item Intervals of equal width.
  \item Intervals containing equal numbers of points.
\end{itemize}

This gives rise to the histogram and spinogram, analogues of the bar chart and spine chart respectively. A long standing tradition is that no gaps are displayed between adjacent rectangles when used for originally continuous data. Two examples are shown in Figure \ref{fig:cont-examples}.

@@

% \subsection{Nested data}
% \label{sub:nested_data}
% 
% Use space-filling!
% 
% We distinguish nested data from crossed data by the number of interactions with missing values.  Completely crossed data contains every combination of all variables.  Nested data contains 
% 
% 
% Aka sparse vs. dense.
% 
% It is not always possible to determine nested data from inspection of the data alone, but may need knowledge of the data collection.  For example, if a survey of teachers within schools was performed, and each teacher was identified by a identifier unique within a school, it might look like the data was only missing a few combinations, when in fact it is meaningless.  
% 
% 
% 
% Or when you have a hierarchy.
% 
% Example dataset - Doug Bates teachers/schools?
% 
% Possibilities: create new variable which is interaction of nested vars, or drop missings/zeros.  Treemaps deal almost exclusively with nested data.
